Defficiency Screen Functional Analysis

### 2025-08-10 DeficiencyScreen/Prelim_Analysis.Rmd
ana update

library(org.Dm.eg.db)
##library(goseq)
library(rrvgo)
library(plotly)
```

@@

##DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))
##colnames(DfKey)[c(1,3)] = c("BDSC", "FBab")
##DfKey$FollowUp = "FALSE"
##DfKey$Excluded = "FALSE"

#old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names)

@@

fullDfKey = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250613_FullDfData.csv"),row.names = 1)

@@

py.only.FBgn = c("lncRNA:CR46532", "FBgn0288728", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

@@

py.only.ens = c("lncRNA:CR46532", "asRNA:PS4", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.CHR = c("lncRNA:CR46532", "3R", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.Start = c("lncRNA:CR46532", 16720127, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.End = c("lncRNA:CR46532", 16729959, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.Strand = c("lncRNA:CR46532", 1, "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

add.gene.locations = data.frame(FBgn = py.only.FBgn,
                                Symbol = py.only.sym,
                                ensembl = py.only.ens,
                                CHR = py.only.CHR,
                                Start = py.only.Start,
                                End = py.only.End,
                                Strand = py.only.Strand)

@@

newdfonly = setdiff(newdf, olddf)
olddfonly = setdiff(olddf, newdf)

newfbonly = setdiff(newfb, oldfb)
oldfbonly = setdiff(oldfb, newfb)

newsymonly = setdiff(newsym, oldsym)
oldsymonly = setdiff(oldsym, newsym)

fbinnewdfonly = unique(fulldf[fulldf$Deficiency %in% newdfonly, "FBgn"])
fbinolddfonly = unique(unlist(genesinolddf[olddfonly]))

@@

@@

```{r echo = F}
##kegg = readRDS(gzcon(url(paste0(git.dir,"DeficiencyScreen/DataFiles/kegg.symbol2path.RData"))))
##kegg.names = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/KEGG.names.csv"), row.names = 1)

Moderate.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/ModerateGOAnalysis.csv"), row.names = 1)
Strong.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/StrongGOAnalysis.csv"), row.names = 1)
Repress.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/RepressorGOAnalysis.csv"), row.names = 1)
Lethal.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/LethalGOAnalysis.csv"), row.names = 1)
Enhancer.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/EnhancerGOAnalysis.csv"), row.names = 1)

genesingo=as.list(org.Dm.egGO2ALLEGS)

##dfbar = data.frame(Enhancer = c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),
##                   NoEffect = c(DfKey[DfKey$Modifier == "No Effect",], dfdata[dfdata$Modifier == "No Effect",]),
##                   Lethal = c(DfKey[DfKey$Modifier == "Lethal",], dfdata[dfdata$Modifier == "Lethal",]),
##                   c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),
##                   Moderate = c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),)

#old = hist(as.numeric(DfKey$MLE))
#new = hist(as.numeric(dfdata$MLE))

```{r echo = F, warning=F}

sigs = unique(fulldf[fulldf$Modifier == "Enhancer", "Symbol"])
genes = setNames(rep(0, length(unique(fulldf$Symbol))), unique(fulldf$Symbol))
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
Enhancer.GO = GO.wall
##write.csv(Suppressor.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/SupressorGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Repress.KEGG = KEGG
##write.csv(Repress.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/RepressorKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Moderate == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Moderate.GO = GO.wall
##write.csv(Moderate.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/ModerateGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Moderate.KEGG = KEGG
##write.csv(Moderate.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/ModerateKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Strong == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Strong.GO = GO.wall
##write.csv(Strong.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/StrongGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Strong.KEGG = KEGG
##write.csv(Strong.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/StrongKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Lethal == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Lethal.GO = GO.wall
##write.csv(Lethal.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/LethalGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Lethal.KEGG = KEGG
##write.csv(Lethal.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/LethalKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Enhancer == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Enhance.GO = GO.wall
##write.csv(Enhance.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/EnhancerGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Enhance.KEGG = KEGG
##write.csv(Enhance.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/EnhancerKEGGAnalysis.csv")

@@

GO.data = Enhancer.GO
GO.ontology = "BP"
GOBP.fig = GO20(GO.data, GO.ontology)

@@

>>>>

library(org.Dm.eg.db)
library(goseq)
library(rrvgo)
library(plotly)
library(VennDiagram)
genesingo=as.list(org.Dm.egGO2ALLEGS)

@@

DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))
##colnames(DfKey)[c(1,3)] = c("BDSC", "FBab")
##DfKey$FollowUp = "FALSE"
##DfKey$Excluded = "FALSE"

old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names)

@@

fullDfKey = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250618_FullDfData.csv"),row.names = 1)

@@

py.only.FBgn = c("FBgn0289441", "FBgn0288728", "FBgn0289439", "FBgn0289444", "FBgn0289438", "FBgn0289443")

@@

py.only.ens = c("86703978", "86703974", "86703976", "86703981", "86703975", "86703980")

py.only.CHR = c("3R", "3R", "3R", "2L", "2L", "2R")

py.only.Start = c(7341783, 16720127, 13056818, 7595504, 9747046, 20149996)

py.only.End = c(7344884, 16729959, 13057676, 7597500, 9747785, 20150499)

py.only.Strand = c(-1, 1, -1, -1, -1, -1)

add.gene.locations = data.frame(FBgn = py.only.FBgn,
                                Symbol = py.only.sym,
                                ensembl = py.only.ens,
                                CHR = py.only.CHR,
                                Start = py.only.Start,
                                End = py.only.End,
                                Strand = py.only.Strand)
gene.location = rbind(gene.location, add.gene.locations)

## redo genes in each Df
gl = apply(gene.location[,c("Start", "End")], MARGIN = 1, min)
gr = apply(gene.location[,c("Start", "End")], MARGIN = 1, max)
Dfl = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, min)
Dfr = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, max)

  if(nrow(temp) > 0){
    temp$Df = fullDfKey$Deficiency[i]
    genesindf = rbind(genesindf, temp[, c("Df", "FBgn", "Symbol")])
  }
  i=i+1
}

```{r echo = F}
#all the current Df data after Jonah omissions
#fullDfKey

##genes in the current Df data
##genesindf

##the old Df data before Jonah omissions
DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))

##Genes that were originally associated with the old Df data
old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)

@@

  ##Sample titles as strings. Only fill in up to your number of selected categories
  set1 = "All old Df"
  set2 = "All new Df"
  set3 = "Old Suppressors/Strong Suppressors"
  set4 = "New Suppressors"

  ##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
  s1 = DfKey$FBab..
  s2 = fullDfKey$FBab[fullDfKey$Omit == F]
  s3 = DfKey$FBab..[grep("Sup",DfKey$Modifier)]
  s4 = fullDfKey[fullDfKey$Omit == F,]
  s4 = s4$FBab[grep("Sup", s4$Modifier)]
  main = ""

@@

trimmed.Df = fullDfKey[fullDfKey$Omit == "FALSE",]
trimmed.genes = genesindf[genesindf$Df %in% trimmed.Df$Deficiency,]
#trimmed.sup = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$MLE >= 2,] 
trimmed.sup = fullDfKey[(fullDfKey$Omit == "FALSE" & fullDfKey$MLE == 2) &
                         fullDfKey$Modifier ==  "No Effect",] 
#trimmed.sup = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$Modifier ==  "Suppressor",] 
trimmed.supgenes = genesindf[genesindf$Df %in% trimmed.sup$Deficiency,]

sigs = unique(trimmed.supgenes$Symbol)
genes = setNames(rep(0, length(unique(trimmed.genes$Symbol))), unique(trimmed.genes$Symbol))

#randoms = sample(1:length(genes), 6061)

genes[sigs] = 1
#genes[randoms] = 1
pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category

#suppressor.GO = GO.wall
#twodayplus.GO = GO.wall
#twotothree.GO = GO.wall

```{r echo = F, warning=F, message=F}
##GO analysis of all Df with MLE 3 or greater, omitting Dfs with escapers
sigs.Df = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$MLE > 3,]
sigs = genesindf[genesindf$Df %in% sigs.Df$Deficiency, "Symbol"]

genes.Df = fullDfKey[fullDfKey$Omit == "FALSE",]
genes = genesindf[genesindf$Df %in% genes.Df$Deficiency, "Symbol"]
sigs = unique(trimmed.supgenes$Symbol)
genes = setNames(rep(0, length(unique(genes))), unique(genes))

genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category

sup.GO = GO.wall

df.genesingo = gene.location[gene.location$ensembl %in% intersect(gene.location$ensembl, as.character(genesingo[["GO:0030527"]])),]

summary(factor(genesindf[genesindf$FBgn %in% df.genesingo$FBgn, "Df"], levels = unique(genesindf[genesindf$FBgn %in% df.genesingo$FBgn, "Df"])))

fullDfKey[fullDfKey$Deficiency == "Df(2L)ED1473",]

```{r echo = F, warning=F, include = F}

@@

@@

GO.data = sup.GO
GO.ontology = "BP"
GOBP.fig = GO20(GO.data, GO.ontology)

@@

sig.sup.GOcats = sup.GO[sup.GO$adjp <= .05, "category"]

ss.gocat.genes = genesingo[sig.sup.GOcats]

ss.gocat1.genes = (gene.location[gene.location$ensembl %in% as.character(ss.gocat.genes[[1]]),])

ss.gocat1.dfs = genesindf[genesindf$FBgn %in% ss.gocat1.genes$FBgn,]

(unique(ss.gocat1.dfs$Df))
